Taking Dyads Seriously

### 2018-02-09 paper/4_app.tex
more changes to ap section

These studies were also selected based on how recent they were and whether they had more then 100 citations.\footnote{Note that we chose papers with at least 100 citations as an indicator of influence within the discipline. Of course, by the criteria of influence many other papers could have been considered for the replication. However, as described in the following sections, the AME model has specific data requirements, which limited the scope of potential papers.}  Each of these pieces, published in prominent journals well-known in their respective literatures, posited a theory in which interdependencies are consequential. However, reflecting the dominant approach in the literature, each of the authors tested their hypothesis by employing some form of a general linearized regression. These GLM models ignore dyadic interdependencies, except as they may reveal themselves in included variables, and then adjusting the standard errors in an attempt to account for the clustering of observations.

In the following section, we present two sets of results for each study, one that follows the original study in our re-estimation of the original models and another that tests the central hypothesis within an AME implementation. We demonstrate that AME allows scholars the ability to test their hypotheses of interest, while directly accounting for both theorized and observed interdependences in the data.

@@

We obtained the data for each of these studies from their replication archives and replicated the main results of each article.\footnote{Without exception this was straightforward to accomplish, thanks to an increasing norm in the social sciences of open data sharing.}  It is worth noting that the AME is applicable only where a full set of dyads for an outcome of interest is observable---for example it is unsuitable for studying onsets, where dyad-years representing conflict continuations are removed from the sample. The AME is also not applicable to analyses of case-level data, for example, studies that examine the decision to go to war by particular states.

% SM: maybe too on the nose (Whether putative control variables, for example, are precisely estimated is less important.) 
Typically, scholars are concerned with conducting hypothesis testing on a small set of independent variables that they have assembled in a complicated empirical model. For example, in the study by Reiter and Stam (2003) the goal is to determine whether in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---it is the democratic country or the personalist country that is associated with a greater likelihood of a dispute dyad being in a dispute.  In a regression including fourteen covariates, they focus on whether the democratic nature of the initiating country is a statistically significant covariate.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framework. Each of the other four studies also has a crucial finding that we hone in on to further draw into focus the potential differences brought about by utilizing an AME estimation procedure.  In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{does not replicate} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.\footnote{Full tabular results for each of the original and reestimated models are presented in the Appendix.}

\begin{table}[ht]
\centering
\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network-based approach.}
	\begin{tabular}{l p{7cm} l} \toprule
		\multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Replicates \\ 
			& &  in a Network Model? \\ \toprule
		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Partially Replicates} \\ \midrule
		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}\\ \midrule
		Rose (2004) & WTO Membership Does not Affect Trade & {Partially Replicates}\\ \midrule
		Weeks (2012) & Bosses, Juntas, and Strongmen are more Aggressive, Machines are Not & {Does Not Replicate} \\\midrule
		Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Does Not Replicate}\\ \bottomrule
	\end{tabular}
	\label{tab:modelFindingSumm}
\end{table}

An important takeaway here is that many studies make knowledge claims based on the statistical significance of a small set of covariates, or the differences between these covariates. These differences may change dramatically when interdependencies are taken into account directly, as they do in the case of this study. Beyond just comparing parameter estimates, we examine how well each approach can represent the data generating process using an out-of-sample cross validation strategy. Specifically, for each study, we randomly divide the data into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.
Then for each $s \in \{1,\ldots,k\}$, we:

>>>>

These studies were also selected based on how recent they were and
whether they had more then 100 citations.\footnote{Note that we chose
  papers with at least 100 citations as an indicator of influence
  within the discipline. Of course, by the criteria of influence many
  other papers could have been considered for the
  replication. However, as described in the following sections, the
  AME model has specific data requirements, which limited the scope of
  potential papers.}  Each of these pieces, published in prominent
journals well-known in their respective literatures, posited a theory
in which interdependencies are consequential. However, reflecting the
dominant approach in the literature, each of the authors tested their
hypothesis by employing some form of a general linearized
regression.

In the following section, we present two sets of results for each
study, one that follows the original study in our re-estimation of the
original models and another that tests the central hypothesis within
an AME implementation. We demonstrate that AME allows scholars the
ability to test their hypotheses of interest, while directly
accounting for both theorized and observed interdependencies in the
data. Directly accounting for interdependences can change the outcome
of regression models, both in substance and in estimates. In part, this outcome
follows from AME's ability to precisely capture higher-order
interdependencies, whereas GLM approaches implicitly ignore dyadic
interdependencies. As this is a widely-known limitation of GLM
approaches, scholars often attempt to account for clustering of
observations by including additional variables and adjusting the
standard errors of the resulting estimates. At best, this method
introduces noise and imprecision into results, and at worst can
produce misleading outcomes.

@@

We obtained the data for each of these studies from their replication
archives and replicated the main results of each
article.\footnote{Without exception this was straightforward to
  accomplish, thanks to the authors'  and an increasing norm in the social sciences of
  open data sharing.}  It is worth noting that the AME is applicable
only where a full set of dyads for an outcome of interest is
observable---for example it is unsuitable for studying onsets, where
dyad-years representing conflict continuations are removed from the
sample. The AME is also not applicable to analyses of case-level data,
for example, studies that examine the decision to go to war by
particular states.

% SM: maybe too on the nose (Whether putative control variables, for
% example, are precisely estimated is less important.) 
%% MJF: 2/9-- took this part out, because I don't think that it's necessar.
% Typically, scholars are concerned with conducting hypothesis testing on a small set of independent variables that they have assembled in a complicated empirical model. For example, in the study by Reiter and Stam (2003) the goal is to determine whether in mixed dyads---consisting of a democratic country and country ruled by a personalist dictatorship---it is the democratic country or the personalist country that is associated with a greater likelihood of a dispute dyad being in a dispute.  In a regression including fourteen covariates, they focus on whether the democratic nature of the initiating country is a statistically significant covariate.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framewo
%rk.
Each of the five studies has a crucial finding that we hone in on to
further draw into focus the analytical power of the AME estimation
procedure.  In Table~\ref{tab:modelFindingSumm}, we present the
overall results; the term \textit{Remains after accounting for
  dependencies?} indicates only that the sign and/or significance of
the putatively crucial finding in the original study is not found to
hold in the AME estimation.\footnote{Full tabular results for each of
  the original and reestimated models are presented in the Appendix.}

\begin{table}[ht]
\centering
\caption{Here we provide a brief summary of the key variable in each
  of the five replications and a note about whether or not the
  highlighted finding remains when using our network-based approach.}
	\begin{tabular}{l p{7cm} l} \toprule
		\multirow{2}{*}{Study} & \multirow{2}{*}{Central
                  Finding} &  Remains after \\ 
			& &  accounting for dependencies? \\ \toprule
		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Partially Replicates} \\ \midrule
		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}\\ \midrule
		Rose (2004) & WTO Membership Does not Affect Trade & {Partially Replicates}\\ \midrule
		Weeks (2012) & Bosses, Juntas, and Strongmen are more Aggressive, Machines are Not & {Does Not Replicate} \\\midrule
		Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Does Not Replicate}\\ \bottomrule
	\end{tabular}
	\label{tab:modelFindingSumm}
\end{table}

An important takeaway here is that as a result of the incompatibility
between GLM models and the underlying processes that shape
international events, many scholars are forced to make knowledge claims
based on the statistical significance of a small set of covariates, or
the differences between these covariates. These differences may change
dramatically when interdependencies are taken into account directly,
as they do in the case of this study. Beyond just comparing parameter
estimates, we examine how well each approach can represent the data
generating process using an out-of-sample cross validation
strategy. Specifically, for each study, we randomly divide the data
into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$
is assigned.

Then for each $s \in \{1,\ldots,k\}$, we:
